=======
---
title: "Lab6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(DALEX2)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")
```
###Individual Part
## Kelsey McKenna
#Question: Which airports had the highest probability of a delayed flight coming into Colorado and what was the degree of this probability?
```{r}
delayplaces <- COflights %>% group_by(ORIGIN) %>%
 filter(ARR_DELAY >15) %>% summarise(Count = n()) 
delayplaces
totalplaces <- COflights %>% group_by(ORIGIN) %>% count()
delayplaces <- delayplaces %>%
  left_join(totalplaces, by = 'ORIGIN')
delayplaces <- delayplaces %>%
  mutate(Total = Count/n) %>% select(ORIGIN, Count, Total) %>% 
  arrange(desc(Total)) 
ggplot(data = delayplaces, aes(x=Total)) +
  geom_histogram(bins=15)+
  ggtitle('Histogram of Places by Delay')
```
As we can see from the histogram, there is one location that had the most flights by a lot at a .45 percent chance of having a flight delay. This location was in twin fall Idaho. There were clearly a couple more places with very high probabilities of having delayed flights. Amarillo, Texas was next followed by Charleston in South Carolina, and Albany, New York. Shockingly there is no regional correlation between these places and we cannot for example conclude that most delays occur in a certain region of the US. We can't even conclude that all these places are far from Denver, since Idaho is relatively close. Also, we should note that the average probability of having a delay into Denver is around .15 and the four highest airports appear to be somewhat of outliers having probabilities around and above .3.

## James Burke
#Question: Which months in the year had the highest probability of a delayed flight coming into Colorado and what was the degree of this probability?
```{r}
mtdelay <- COflights %>% filter(!is.na(ARR_TIME)) %>% mutate(truDelay =  ARR_DELAY >= 15 )
delay <- COflights %>% filter(!is.na(ARR_TIME)) %>% filter(ARR_DELAY >= 15) 
total <- COflights %>% filter(!is.na(ARR_TIME))
late <- COflights %>% filter(!is.na(ARR_TIME)) %>% select(MONTH, delay) %>% arrange(desc(ARR_DELAY))
ggplot(data = late, aes(x = MONTH, y = )) +
  geom_point() +
  ggtitle('Delays Per Month')
```
## Matthew Copeland
#Question: Which airline carrier has the most delayes at Denver?
```{r}
carriers <- COflights %>% select(CARRIER, ARR_DELAY) %>% arrange(desc(ARR_DELAY))
carriers
ggplot(data = carriers)+
  geom_col(aes(x = CARRIER, y = ARR_DELAY, fill = CARRIER),position = "dodge", na.rm = TRUE)+
  ggtitle('Arrival Delay v.s Carriers')+
  ylab('Total Arrival Delaye Minutes')+
  xlab('Carrier')+
  labs(fill = 'Carrier')
```
As the graph shows american airlines has the longets delays of planes landing in colorado. 

##Shreeya Basrur Individual
Question: How do weather delays contribute to overall delays? Over the course of a year, how does Colorado's variety of weather patterns due to having four distinct seasons impact the number of weather delays?
```{r}
delayedflights <- COflights %>% filter(ARR_DELAY>15) %>% filter(YEAR == "2017")%>% summarise(count = n())
delayedflights

year<-COflights %>% group_by(DEST == "DEN") %>% filter(YEAR == "2017")%>% summarise(count = n())

weatherdelay <-COflights %>% select(WEATHER_DELAY, DAY_OF_MONTH) %>% summarise(count = n())
weatherdelay
prob <-79030/476809
prob

weather <- COflights %>% filter(WEATHER_DELAY>15) %>% summarise(count = n())
weather


ggplot(data = COflights)+
  geom_col(mapping = aes(x = MONTH, y = WEATHER_DELAY),position = "dodge", na.rm = TRUE, fill = "yellow") +
  ggtitle("Average weather delay over 2017 by month")+
    xlab("Months") +
    ylab("Weather Delay")
```

Findings: The probablility of a weather delay given all delays was .165. That means that weather delays played a small role in all the delays that DIA experiences. This probability was derived by finding the total count of planes delayed by 15 minutes and finding the percentage of those planes that were delayed by weather. Plotting that data, January has the most delays while September and November had the fewest delays.

Conclusion: Although weather delays don't contribute that much to the enormity of the problem of delayed flights, the weather of January, March, and August seemed to cause more delayed flights while the opposite effect took place in September and November.


##Team Work
#Quesiton: Do places with a longer security wait also have more delayed flights? 
#Importance: This is important, because it is necessary to see whether security waits are related to delayed flights, if there were a correlation this would indicate that some airports were less efficient than other overall. 
```{r}
securitywait <- COflights %>% group_by(ORIGIN) %>% 
  select(ORIGIN, SECURITY_DELAY) %>% filter(SECURITY_DELAY >15) %>%
  summarise(secure = n())
delaywait <- COflights %>% group_by(ORIGIN) %>% 
  select(ORIGIN, LATE_AIRCRAFT_DELAY)  %>% filter(LATE_AIRCRAFT_DELAY>15) %>%
  summarise(late = n())
delaysecure <- merge(securitywait, delaywait, by="ORIGIN")


delaysecure <- delaysecure %>% filter(ORIGIN != "DEN")
securitywait
delaywait

ggplot(data=delaysecure)+
  geom_jitter(aes(x = secure, y = late))+
  geom_smooth(aes(x = secure, y = late))
  xlab("Security Delays")+
  ylab("Flight Delays")

ggplot(A,aes(x,y)) +geom_point() +geom_point(data=B,colour='red')
```
Findings: As you can see from the plot, (we removed DEN because it was skewing the data and clearly an outlier) there is a positive relationship with Security Delays and Flight delays. In other words, airports that have more security delays are more likely to have flight delays. Therefore, these variables are related. 

##Team Summary
I, Kelsey McKenna, used probabilities to find my individual plot and come to a conclusion on which airports have the most delays. I didn't find a regional correlation between them. I also did the team plot relating security delays to flight delays and found a positive correlation between the two based on the plot. 

I, Shreeya Basrur, created a bar graph to analyse the distribution of weather delays over a single year by month. This was done using the geom_col function, with the NA values taken out and the graph modified to not be stacked. The plot was filled in yellow for contrast against the gray background. In addition, a plot title and axis titles were added for clarity.

>>>>>>> master
